---
title: "Unit-1 Homework"
author: "Steven Garrity"
date: "8/25/2019"
output: html_document
---

### My Data Science Profile
```{r}
labels = c("Data Viz","ML","Math","Stats","CS", "Comms","Domain")
myscores = c(1.5,2.5,1,2,1.5,2,0)
myprofile = data.frame(Category=labels, Score=myscores)
barplot(myprofile$Score, names.arg=myprofile$Category, col='blue', 
        ylab="Score", main="Garrity DS Profile", ylim=c(0,5))
```

### Chi-square population
```{r}
population = rchisq(10000000, df=2)
```

```{r}
hist(population, col='blue',main="Population distribution, N = 10,000,000")
#text(25,3800000,expression(paste(mu, "=")))
text(25,3800000, bquote(mu == .(mean(population))))
text(25,3500000, bquote(sigma == .(sd(population))))
```
#### Control Parameters
```{r}
n = 50 # sample size per sample for 1st distribution
simulations = 10000 #number of xbars we will generate.
```

#### Data Holder
```{r}
xbar_holder1 = numeric(simulations) # This will hold all the sample means for the chi-square distribution.
```

#### Simulate and Store
Generate 10000 samples each of size 50 and find the mean of each sample.  Then store each mean in the xbar_holder vector.
```{r}
for (i in 1:simulations)
{ 
  sample1 = sample(population,n)
  xbar1 = mean(sample1)
  xbar_holder1[i] = xbar1
}
```

## display the distribution of sample means (plot a histogram of the sample means)
```{r}
require(grDevices)

hist(xbar_holder1, col = "blue", main = paste("Distribution of the sample mean: n = ", n), xlab = "Sample Means", xlim = c(0,4))
text(3.25,2250, bquote(bar(x) == .(mean(xbar_holder1))))
text(3.25,2000, bquote(s == .(sd(xbar_holder1))))
```



## summary statistics of the distribution of the simulated sample means. 
```{r}
summary(xbar_holder1) #5 number summary and the mean
sd(xbar_holder1) # standard deviation of distribution 1
```


### Beachcomber Data
```{r}
patron_age = c(25,19,37,29,40,28,31)
patron_mean = mean(patron_age)
patron_s = sd(patron_age)
patron_n = length(patron_age)

lower_bound = patron_mean-(2.447*(patron_s/sqrt(patron_n)))
upper_bound = patron_mean+(2.447*(patron_s/sqrt(patron_n)))
```

```{r}
t.test(patron_age, mu=21, conf.level = 0.95)
```
