---
title: "DDS_Unit2_Homework"
author: "Steven Garrity"
date: "9/1/2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r "ls", echo=FALSE}
system("ls")
```

## Load Data:
```{r playerbb, results='hide'}
playerbb = read.csv("/Users/stevengarrity/SMU_MSDS/DS6306_DoingDataScience/DDS_Git/Unit 2/PlayersBBall.csv", header = TRUE)
summary(playerbb)
```

## Question 1
Use the PlayerBBall.csv dataset to visually represent (summarize) the number of players in each position.

```{r, echo=FALSE}
test <- droplevels(playerbb)
playerbb %>% ggplot(aes(x=position, fill=position)) + geom_bar() + ggtitle("Number of Players in Each Position")

test %>% ggplot(aes(x=position, fill=position)) + geom_bar() + ggtitle("Number of Players in Each Position")
```

## Question 2
Use the dataset to visually investigate the distribution of the weight of centers (C) is greater than the distribution of the weight of forwards (F).

```{r}
# test <- playerbb %>% pull(weight)
# player_weights <- playerbb[(playerbb$position == "C" | playerbb$position == "F"),]
# 
# # the next two plots are identical, however, I can't get a legend to show
# ggplot(playerbb, aes(x=weight)) +
#   geom_histogram(data=subset(playerbb, position == "F"), color = "black", fill = "blue", alpha = 0.7) +
#   geom_histogram(data=subset(playerbb, position == "C"), color = "black", fill = "red", alpha = 0.7)
# 
# ggplot(player_weights, aes(x=weight)) +
#   geom_histogram(data=subset(player_weights, position == "F"), color = "black", fill = "blue", alpha = 0.7) +
#   geom_histogram(data=subset(player_weights, position == "C"), color = "black", fill = "red", alpha = 0.7)

# try something different
weight_C <- droplevels(subset(playerbb, position == "C", select=c(weight,position)))
weight_F <- droplevels(subset(playerbb, position == "F", select=c(weight,position)))
weights_C_F = rbind(weight_C,weight_F)
rm(weight_C)
rm(weight_F)

# weights_C_F %>% ggplot(aes(x=weight, color=position)) +  geom_histogram()


#####
# ggplot(weights_C_F, aes(x=weight, fill = position)) + 
#   geom_histogram(alpha = 0.5) + 
#   scale_fill_manual(name="position",values=c("red","darkgray"),labels=c("C","F"))

ggplot(weights_C_F,aes(x=weight))+ 
    geom_histogram(data=subset(weights_C_F,position=='C'),aes(fill=position),alpha=0.6)+
    geom_histogram(data=subset(weights_C_F,position=='F'),aes(fill=position),alpha=0.6)+
    scale_fill_manual(name="position", values=c("red","gray35"),labels=c("C","F"))+
  ggtitle("Weight distribution for Centers and Forwards")+
  xlab("weight (lbs)")
  

```

# Question 3
Use the dataset to visually investigate if the distribution of the height of centers (C) is greater than the distribution of the height of forwards (F).


```{r}
## for height-related questions:
temp_var <- str_split_fixed(playerbb$height,n = 2,pattern = "-")
playerbb$feet_height = as.numeric(temp_var[,1])+(as.numeric(temp_var[,2])/12)
playerbb$inches_height = (as.numeric(temp_var[,1])*12)+as.numeric(temp_var[,2])

ggplot(playerbb,aes(x=feet_height))+ 
    geom_histogram(data=subset(playerbb,position=='C'),aes(fill=position),alpha=0.6)+
    geom_histogram(data=subset(playerbb,position=='F'),aes(fill=position),alpha=0.6)+
    scale_fill_manual(name="position", values=c("red","gray35"),labels=c("C","F"))+
  ggtitle("Height distribution of Centers and Forwards")+
  xlab("height (ft)")

```

# Question 4
Use the dataset to visually investigate if the distribution of height is different between any of the positions.

```{r}
playerbb %>% ggplot(aes(x = position, y=feet_height, fill=position)) + geom_boxplot() +
  ylab("height (ft)") + ggtitle("Height Distribution by Position")

```

# Question 5
Use the dataset to investigate how the player’s height is related to the player’s weight. How does height change as the weight changes?

```{r}
library(gridExtra)

p1 <- ggplot(data = playerbb, mapping = aes(x=feet_height, y=weight)) +
  geom_point() +
  geom_smooth() +
  xlab("height (ft)") +
  ylab("weight (lbs)")

p2 <- ggplot(data = playerbb, mapping = aes(x = feet_height, y = weight, linetype = position, color = position)) + 
geom_point() +
geom_smooth() +
  xlab("height (ft)") +
  ylab("weight (lbs)")

grid.arrange(p1, p2, nrow = 1)
```

# Question 6
Is their any difference in the relationship between height and weight between positions?  Are height and weight related differently for different positions.

```{r}
ggplot(data = playerbb) + geom_point(mapping = aes(x = feet_height, y = weight, color = position)) + geom_smooth(mapping = aes(x = feet_height, y = weight)) + facet_wrap(~position) +
  ylab("weight (lbs)") +
  xlab("height (ft)") +
  ggtitle("Height v Weight by Position")
```

# Question 7
A historian would like to investigate the claim that the heights of players have increased over the years.  Analyze this claim graphically / visually.

```{r}
# ggplot(data = playerbb) + geom_point(mapping = aes(x = year_start, y = feet_height))

playerbb %>% ggplot(aes(x = factor(year_start), y = feet_height)) + geom_boxplot(color = "dodgerblue4") +
  theme(axis.text.x = element_text(angle = 70)) +
  ylab("height (ft)") +
  xlab("Starting Year") +
  ggtitle("Height by Starting Year")
```

# Question 8
Create a 3D plot of height vs. weight vs. year and color code the points by position.

```{r}
library(ggthemes)
library(plotly)

p <- plot_ly(playerbb, x = ~feet_height, y = ~weight, z = ~year_start, color = ~position) %>%
   add_markers() %>%
   layout(scene = list(xaxis = list(title = 'height (ft)'),
                      yaxis = list(title = 'weight (lbs)'),
                      zaxis = list(title = 'starting year')))
p
```

# Question 9
Go to this website and use one of the 50 best plots to visualize some aspect of the data and provide at least one insight.  You will present your work in breakout! http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html

```{r}
#do shorter players have shorter careers than taller players
# playerbb %>%  ggplot(aes(x=feet_height, y=(year_end-year_start))) +
#   geom_point()

# test <- droplevels(playerbb)

# library(treemapify)
# 
# 
# months = unique(playerbb$birth_month)
# number_born = c()
# 
# for(i in 1:length(months)-1)
# {
#   number_born[i] = sum(with(playerbb, birth_month==months[i]))
# }
# hist(number_born)
# 
# months_filtered = months[1:12]
# dfborn = data.frame(number = number_born, monthsdf = months_filtered)
# 
# 
# 
# 
# dfborn$monthsR <- factor(dfborn$monthsdf, c("January","February","March","
#                                           April","May","June","July","August","September","October","November","December"))
# dfborn %>% ggplot(aes(x=monthsR, y=number)) + geom_bar(stat="identity") +
#   theme(axis.text.x = element_text(angle = 70))
# 
# dfborn %>% ggplot(aes(area = number, fill = monthsR)) +
#   geom_treemap() 

playerbb$careerlength = playerbb$year_end - playerbb$year_start

playerbb %>% ggplot(aes(x=position, y=careerlength)) +  geom_violin(color="darkgray") +  
labs(title="Violin plot", 
       subtitle="Career Length by Position",
       caption="Source: Playersbball.csv",
       x="position",
       y="career length (years)")

```

# Question 10
Visually test the claim that the distribution of incomes increase (mean or median) as the education level rises.

```{r}
edincome = read.csv("/Users/stevengarrity/SMU_MSDS/DS6306_DoingDataScience/DDS_Git/Unit 2/Education_Income.csv", header = TRUE)
# summary(edincome)

edincome$EducR <- factor(edincome$Educ, c("<12","12","13-15","16",">16"))


edincome %>% ggplot(aes(x = EducR, y = Income2005)) + geom_boxplot(color = "dodgerblue2", fill = "gray20") + ylab("Income ($)") +
  xlab("Years of Education") +
  ggtitle("Income by Years of Education") +
  ylim(0,250000)
```




