---
title: "Unit3_Homework"
author: "Steven Garrity"
date: "9/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages:
require(tidyverse)
require(GGally)
```

# DS News of the Week
https://www.nature.com/articles/s41586-019-1335-8

article: https://newscenter.lbl.gov/2019/07/03/machine-learning-algorithms-can-uncover-hidden-scientific-knowledge/

https://www.iflscience.com/technology/artificial-intelligence-set-loose-on-old-scientific-papers-discovers-something-humans-missed/

# Part 1A
Using the FIFA player data set, filter the data set to create a dataframe that has just the Left Midfielders (LM) and Left Forwards (LF). Use Ggally and ggpairs() and the dataset you created above above, to plot the categorical variable Position (LM and LF), versus the continuous variables Acceleration and Agility. Given the plot above, what relationships do you see?  Comment on these.
```{r}
# require(GGally)
fifa = read.csv("/Users/stevengarrity/SMU_MSDS/DS6306_DoingDataScience/DDS_Git/Unit 3/FIFA Players.csv", header = TRUE)
# head(fifa)

### look at the number of records per position from the native data:
fifa %>% group_by(Position) %>% summarize(count = n()) %>% arrange(count) %>% print(n=30)

### create new data frame that only contains filtered data:
q1_fifa <- fifa %>%
filter(Position == "LM" | Position == "LF") %>%
select(Acceleration, Agility, Position) 

###clean up the position factor
q1_fifa$posFact = factor(q1_fifa$Position, levels=c("LF","LM"))

### sanity check...make sure we have the data that we think we do:
q1_fifa %>% group_by(posFact) %>% summarize(count = n())

### EDA plotting of the data:
# q1_fifa$posFact <- factor(q1_fifa$posFact, levels=rev(levels(q1_fifa$posFact)))
q1_fifa %>% select(Acceleration, Agility, posFact) %>% ggpairs(aes(color=posFact, alpha=0.2)) + ggtitle("LM vs LF EDA")
```

# Part2B
Your client would like to formally test if the mean agility rating of left midfielders is different than that of the left forwards.  Perform a 6 – step t-test to test for the difference in these means.  (You may skip step 2 (draw and shade) if you like.  If you are unfamiliar with the 6-step hypothesis test, see Stat 1 slides or the Bridge Course to review the 6-step hypothesis test.)
```{r}
# the density plots and boxplots show that the assumption of normally distributed samples appears to be valid (there is some left skew in both distributions, so we should use some caution in interpretation)
# now let's test for equal variances:
var.test(q1_fifa$Agility ~ q1_fifa$posFa, alternative="two.sided")

###################
#draw and shade
shade <- function(df, alpha, h0 = 0, sides='both', t_calc=NULL) {
  e_alpha = alpha
  if(sides == 'both'){
    e_alpha = alpha / 2
  }
  cv = abs(qt(e_alpha, df))
  curve(dt(x, df), from = -4, to = 4, ylab='P(x)', xaxt='n') 
  abline(v = 0, col = "black", lwd = 0.5)
  labels = h0
  at = 0
  if(sides == 'both' | sides == 'left'){
    x <- seq(-4, -abs(cv), len = 100) 
    y <- dt(x, df)
    polygon(c(x, -abs(cv)), c(y, min(y)), col = "blue", border = NA)
    lines(c(-cv, -cv), c(0, dt(-cv, df)), col = "black", lwd = 1)
    text(-cv - (4 - cv) / 2, 0.05, e_alpha)
    labels = c(round(-cv, 3), labels)
    at = c(-cv, at)
  }
  if(sides == 'both' | sides == 'right'){
    x <- seq(abs(cv), 4, len = 100)
    y <- dt(x, df)
    polygon(c(abs(cv), x), c(min(y), y), col = "blue", border = NA)
    lines(c(cv, cv), c(0, dt(cv, df)), col = "black", lwd = 1)
    text(cv + (4 - cv) / 2, 0.05, e_alpha)
    labels = c(labels, round(cv, 3))
    at = c(at, cv)
  }
  if(is.numeric(t_calc)){
    abline(v = t_calc, col = "red", lwd = 2)
    text(t_calc + 0.5, 0.2, t_calc, col = "red")
  }
  axis(1, at=at, labels=labels)
}
#The above defines the function shade. To use it, you must call it. More examples are in the comments above.
# shade(49, 0.05, 0, t_calc=1.1)

# finally, run a two-sample t-test for equal means
t.test(q1_fifa$Agility ~ q1_fifa$posFact, alternative="two.sided", paired=FALSE, var.equal=TRUE, conf.level = 0.95)

# plot "draw and shade"
shade(1108, 0.05, 0, t_calc=1.8109)
```

# Part 2C
test assumptions
```{r}
## test assumptions:
fifa %>% filter(Position == "LM" | Position == "LF") %>% ggplot(aes(x = Agility, fill = Position)) + geom_density() + facet_wrap(~Position) + 
  labs(title="Agility Kernel Densities: LM vs. LF", x="Agility", y="density") +
  xlim(30,110)

fifa %>% filter(Position == "LM" | Position == "LF") %>% ggplot(aes(x = Position, y = Agility, fill = Position)) + geom_boxplot() + 
  labs(title="Agility Boxplots: LM vs. LF")

fifa %>% filter(Position == "LM" | Position == "LF") %>% 
  ggplot(aes(sample = Agility, color=Position)) + 
  stat_qq() + stat_qq_line() +
  facet_wrap(~Position)
  labs(title="QQ Plots: LM vs. LF")
```

# Part 2
```{r}
fifa %>% group_by(Position) %>% summarize(count = n()) %>% arrange(count) %>% print(n=30)
# we are only looking at the Striker position, so create new data frame
q2_fifa <- fifa %>% filter(Position == "CF" | Position == "CM" | Position == "CB")
q2_fifa$posFact = factor(q2_fifa$Position, levels=c("CF","CM","CB"))
# q2_fifa <- fifa %>% filter(Position == "CAM" | Position == "CDM")
# q2_fifa$posFact = factor(q2_fifa$Position, levels=c("CAM","CDM"))

mental_names <- c("Positioning","Vision","Composure","Interceptions","Aggression")
technical_names <- c("Finishing","LongShots","Penalties","ShotPower","Volleys",
                     "Crossing","Curve","FKAccuracy","LongPassing","ShortPassing",
                     "BallControl","Dribbling","HeadingAccuracy","Marking",
                     "SlidingTackle","StandingTackle")
physical_names <- c("Acceleration","SprintSpeed","Agility","Balance","Reactions",
                    "Jumping","Stamina","Strength")
q2_fifa$mental <- rowMeans(q2_fifa[,mental_names], na.rm=TRUE)
q2_fifa$technical <- rowMeans(q2_fifa[,technical_names], na.rm=TRUE)
q2_fifa$physical <- rowMeans(q2_fifa[,physical_names], na.rm=TRUE)

# convert rating to numeric and then to factor
q2_fifa$rating <- as.numeric(q2_fifa$ST)
q2_fifa$ratingClass = cut(q2_fifa$rating, breaks = c(0,40,60,100), labels = c("Low","Medium","High"))

q2_fifa$ageClass = cut(q2_fifa$Age, breaks = c(0,21,27,33,39), 
                       labels = c("Y","YM","MO","O"))

q2_fifa %>% select(mental, technical, physical, ratingClass, ageClass, posFact) %>%  ggpairs(aes(color=posFact, alpha=0.5))


#### Look at linear correlations
q2_fifa %>% ggplot(aes(x=physical, y=rating, color=posFact)) + geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  labs(x="Physical Rating", y="Overall Player Rating")
q2_fifa %>% ggplot(aes(x=technical, y=rating, color=posFact)) + geom_point() +
   geom_smooth(method="lm", se=FALSE) +
  labs(x="Technical Rating", y="Overall Player Rating")
q2_fifa %>% ggplot(aes(x=mental, y=rating, color=posFact)) + geom_point() +
   geom_smooth(method="lm", se=FALSE) +
  labs(x="Mental Rating", y="Overall Player Rating")
```

# Takeaways & Questions
I'm confused by the factor vs. arrange. Especially why we had to be explicit for the graph to turn out as desired. When we created the data frame with "arrange" it seemed to be ordered correctly, but the plot didn't show as intended. It wasn't until we invoked "order" that things worked as expected.

